---
title: "Assigning country codes to Darwin Core records"
author: "Pieter Provoost"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

In this notebook we are going to assign country codes to Darwin Core records using a shapefile from [Marine Regions](https://www.marineregions.org/downloads.php).

```{r setup, include = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

Load packages:

```{r message = FALSE}
library(dplyr)
library(sf)
library(ggplot2)
library(tibble)
```

Let's read the shapefile. The shapefile is not included in the code repository but can be downloaded from [Marine Regions](https://www.marineregions.org/downloads.php) (see "Marine and land zones: the union of world country boundaries and EEZ's").

```{r message = FALSE}
shapes <- st_read("EEZ_land_union_v3_202003/EEZ_Land_v3_202030.shp", quiet = TRUE)
head(shapes)
```

Now let's generate some random coordinates and convert them to an `sf` object:

```{r message = FALSE}
n_points <- 30000
points <- data.frame(decimalLongitude = runif(n_points, min = -180, max = 180), decimalLatitude = runif(n_points, min = -90, max = 90)) %>%
  st_as_sf(coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
head(points)
```

Now we can join the points and shapes layers using `st_join()`. I'm also adding a column with the original row numbers here, this is necessary because the resulting dataframe `joined` may be larger than the original `points` dataframe due to overlapping polygons.

```{r message = FALSE}
joined <- points %>%
  rownames_to_column(var = "rowname") %>%
  st_join(shapes, join = st_within)
head(joined)
```

Now we can merge the duplicated lines and concatenate country codes (column `ISO_TER1`) where necessary.

```{r message = FALSE}
result <- joined %>%
  group_by(rowname) %>%
  summarize(country_code = paste0(na.omit(unique(ISO_TER1)), collapse = ";")) %>%
  mutate(country_code = na_if(country_code, "")) %>%
  arrange(as.numeric(rowname)) %>%
  st_set_geometry(NULL)
head(result)
table(result$country_code)
```

Finally I'm adding the country codes to the original table:

```{r message = FALSE}
points$country_code <- result$country_code

points %>%
  filter(!is.na(country_code)) %>%
  ggplot() +
    geom_sf(aes(color = country_code), size = 0.3) +
    theme(legend.position = "none")
```

